---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# rTRNG: R package providing access and examples to TRNG C++ library


[![Travis-CI Build Status](https://travis-ci.org/miraisolutions/rTRNG.svg?branch=master)](https://travis-ci.org/miraisolutions/rTRNG)

[TRNG](https://numbercrunch.de/trng/) (Tina's Random Number Generator) is a state
of the art C++ pseudo-random number generator library for sequential and
parallel Monte Carlo simulations. It provides a variety of random number engines
(pseudo-random number generators) and distributions. In particular, *parallel*
random number engines provided by TRNG can be manipulated by `jump` and `split`
operations and hence are suitable to parallel algorithms.

Package **rTRNG** provides the R users with access to the functionality of the
underlying TRNG C++ library, both in R and as part of other projects combining R
with C++.

## Installation

The package can be installed from our 
[github repository](https://github.com/miraisolutions/rTRNG) with:
```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("miraisolutions/rTRNG")
# in order to also build the vignettes, you'll have to run below instead
devtools::install_github("miraisolutions/rTRNG", build_vignettes = TRUE)
```

## Build note

If you try to build the package yourself from source and run `Rcpp::compileAttributes()` during the process, you
need to use a version of **Rcpp >= 0.12.11.2**. Earlier versions like 0.12.11 will not generate the desired
`_rcpp_module_boot_trng` symbol in *RcppExports.cpp*.
