---
output: github_document
################################################################################
# NOTE: 
# Make sure you have added an executable pre-commit hook in your local checkout 
#   .git/hooks/pre-commit
# to prevent accidental commits of README.Rmd not re-knitted to README.md.
# If not, you can copy the hook from inst/tools
#   $ cp inst/tools/pre-commit .git/hooks
################################################################################
---

<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  R.options = list(digits = 6, width = 60) # fit 5 elements per row
)
```
```{r, check-vignette, include = FALSE}
# make sure a vignette with a given name exists, you may have to run
# devtools::install(build_vignettes = TRUE) in case of errors
stop_if_no_vignette <- function(topic, package = "rTRNG") {
  invisible( # prevents opening the vignette
    withCallingHandlers(
      vignette(topic, package),
      warning = function(w) stop(w$message)
    )
  )
}
```

<img src="man/figures/rTRNG.svg" align="right" width="15%" height="15%"/>

# rTRNG: R package providing access and examples to TRNG C++ library

[![CRAN status](http://www.r-pkg.org/badges/version/rTRNG)](https://cran.r-project.org/package=rTRNG)
[![Travis build status](https://travis-ci.com/miraisolutions/rTRNG.svg?branch=master)](https://travis-ci.com/miraisolutions/rTRNG)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/miraisolutions/rTRNG?branch=master&svg=true)](https://ci.appveyor.com/project/riccardoporreca/rtrng)
[![Codecov status](https://img.shields.io/codecov/c/github/miraisolutions/rTRNG/master.svg)](https://codecov.io/github/miraisolutions/rTRNG?branch=master)

**[TRNG](https://numbercrunch.de/trng/)** (Tina's Random Number Generator) is a
state-of-the-art C++ pseudo-random number generator library for sequential and
parallel Monte Carlo simulations. It provides a variety of random number engines
(pseudo-random number generators) and distributions. In particular, *parallel*
random number engines provided by TRNG can be manipulated by `jump` and `split`
operations. These allow to `jump` ahead by an arbitrary number of steps and to
`split` a sequence into any desired sub-sequence(s), thus enabling techniques
such as *block-splitting* and *leapfrogging* suitable to parallel algorithms.

Package **rTRNG** provides the R user with access to the functionality of the
underlying TRNG C++ library, both in R directly or more typically as part of
other projects combining R with C++.

An
[introduction to **rTRNG**](https://user2017.sched.com/event/Axpj/rtrng-advanced-parallel-random-number-generation-in-r)
[[pdf](http://schd.ws/hosted_files/user2017/93/Mirai.rTRNG.useR2017.pdf)]
was given at the useR!2017 conference, and is also available as package
vignette:
```{r check-vignettes, echo=FALSE}
stop_if_no_vignette("rTRNG.useR2017")
stop_if_no_vignette("mcMat")
```
```{r useR2017, eval = FALSE}
vignette("rTRNG.useR2017", "rTRNG")
```

The *sub-matrix simulation* vignette shows **rTRNG** in action for a flexible
*and* consistent (parallel) simulation of a matrix of Monte Carlo variates:
```{r mcMat, eval = FALSE}
vignette("mcMat", "rTRNG")
```

A full applied example of using **rTRNG** for the simulation of credit defaults
was presented at the
[R/Finance 2017](http://past.rinfinance.com/agenda/2017/talk/RiccardoPorreca.pdf)
conference. The underlying code and data are hosted on [GitHub](https://github.com/miraisolutions/PortfolioRiskMC),
as is the corresponding
[R Markdown output](https://rawgit.com/miraisolutions/PortfolioRiskMC/master/RinFinance2017/PortfolioSimAndRiskBig.html).

For more information and references, you can consult the package documentation
page via `help("rTRNG-package")`.


## Installation

You can install the package from CRAN:
```{r cran-install, eval = FALSE}
install.packages("rTRNG")
```

The development version of **rTRNG** can be installed from our 
[GitHub repository](https://github.com/miraisolutions/rTRNG) with:
```{r gh-install, eval = FALSE}
# install.packages("remotes")
remotes::install_github("miraisolutions/rTRNG")
# in order to also build the vignettes, you'll have to run below instead
remotes::install_github("miraisolutions/rTRNG", build_vignettes = TRUE)
```

----

*Build note*

If you try to build the package yourself from source and run
`Rcpp::compileAttributes()` during the process, you need to use a version of
**Rcpp >= 0.12.11.2**. Earlier versions like 0.12.11 will not generate the
desired `_rcpp_module_boot_trng` symbol in *RcppExports.cpp*.


## Examples 

### Use TRNG engines from R

Similar to base-R (`?Random`), **rTRNG** allows to select and manipulate a
*current* TRNG engine of a given *kind* (e.g. yarn2), and to draw from it using
any of the provided `r<dist>_trng` functions:
```{r ex-TRNG.Random}
library(rTRNG)
TRNGkind("yarn2") 
TRNGseed(12358)
runif_trng(15)
```
The special `jump` and `split` operations can be applied to the current
engine in a similar way:
```{r ex-jump.split}
TRNGseed(12358)
TRNGjump(6) # advance by 6 the internal state
TRNGsplit(5, 3) # subsequence: one element every 5 starting from the 3rd
runif_trng(2)
#   => compare to the full sequence above
```

TRNG engines can also be created and manipulated directly as *Reference Class*
objects, and passed as `engine` argument to `r<dist>_trng`:
```{r ex-TRNG.Engine}
rng <- yarn2$new()
rng$seed(12358)
rng$jump(6)
rng$split(5, 3)
runif_trng(2, engine = rng)
```

In addition, parallel generation of random variates can be enabled in
`r<dist>_trng` via `RcppParallel` using argument `parallelGrain > 0`:
```{r ex-parallelGrain}
TRNGseed(12358)
RcppParallel::setThreadOptions(numThreads = 2)
x_parallel <- rnorm_trng(1e5L, parallelGrain = 100L)
TRNGseed(12358)
x_serial <- rnorm_trng(1e5L)
identical(x_serial, x_parallel)
```

### Use TRNG from standalone C++

The TRNG C++ library is made available by **rTRNG** to standalone C++ code
compiled with `Rcpp::sourceCpp` thanks to the `Rcpp::depends` attribute, with
`Rcpp::plugins(cpp11)` enforcing the C++11 standard required by TRNG >= 4.22:

```{Rcpp ex-Rcpp}
// [[Rcpp::depends(rTRNG)]]
// TRNG >= 4.22 requires C++11
// [[Rcpp::plugins(cpp11)]]
#include <Rcpp.h>
#include <trng/yarn2.hpp>
#include <trng/uniform_dist.hpp>
// [[Rcpp::export]]
Rcpp::NumericVector exampleCpp() {
  trng::yarn2 rng(12358);
  rng.jump(6);
  rng.split(5, 2); // 0-based index
  Rcpp::NumericVector x(2);
  trng::uniform_dist<>unif(0, 1);
  for (unsigned int i = 0; i < 2; i++) {
    x[i] = unif(rng);
  }
  return x;
}
```
```{r ex-Rcpp-R}
exampleCpp()
```

### Use TRNG from other R packages

Creating an R package with C++ code using the TRNG library and headers through
**rTRNG** is achieved by

- adding `Imports: rTRNG` and `LinkingTo: rTRNG` to the DESCRIPTION file
- importing one symbol in the NAMESPACE: `importFrom(rTRNG, TRNG.Version)`
- enforcing compilation using C++11 in Makevars[.win] via `CXX_STD = CXX11`
- setting the relevant linker flags in Makevars[.win] via `rTRNG::LdFlags()`
    - Makevars: `PKG_LIBS += $(shell ${R_HOME}/bin/Rscript -e "rTRNG::LdFlags()")`
    - Makevars.win: `PKG_LIBS += $(shell "${R_HOME}/bin${R_ARCH_BIN}/Rscript.exe" -e "rTRNG::LdFlags()")`

### Note about C++ code on macOS

C++ code using the TRNG library (sourced via `Rcpp::sourceCpp` or part of an R
package) might fail on certain systems due to issues with building and linking
against **rTRNG**. This is typically the case for  **macOS**, and can generally
be checked by running

```{r check-linking, eval=FALSE}
rTRNG::check_rTRNG_linking()
```
