---
title: "Flexible and consistent simulation of a matrix of Monte Carlo variates"
author: "Riccardo Porreca, Roland Schmid"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sub-matrix simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
```

```{r include=FALSE, cache=FALSE}
source("utils/read_chunk_wrap.R", echo = FALSE, print.eval = FALSE)
Rcpp::sourceCpp("code/mcMat.cpp", verbose = FALSE, embeddedR = FALSE)
read_chunk_wrap("code/mcMat.cpp")
Rcpp::sourceCpp("code/mcMatParallel.cpp", verbose = FALSE, embeddedR = FALSE)
read_chunk_wrap("code/mcMatParallel.cpp")
read_chunk_wrap("code/mcMat.R")
```

## Consistent sub-simulaton with Rcpp

We use Rcpp to define a function `mcMatRcpp` performing the full sequential Monte Carlo simulation of `nrow` normal i.i.d. samples of `ncol` variables using the `yarn2` generator.
```{Rcpp mcMatRcpp, eval=FALSE}
```
A second function `mcSubMatRcpp` relies on `jump` and `split` operations to perform only a chunk [`startRow`, `endRow`] of simulations for a subset `subCols` of the variables.
```{Rcpp mcSubMatRcpp, eval=FALSE}
```
The parallel-engine nature of the `yarn2` generator ensures the sub-simulation obtained via `mcSubMatRcpp` is consistent with the full sequential simulation.
```{r subMatExampleRcpp}
```
```{r, echo=FALSE}
knitr::kable(cbind.data.frame(M = M, S = S), row.names = TRUE)
```


## Consistent parallel (sub-)simulaton with RcppParallel

We can embed the body of `mcSubMatRcpp` above into a `RcppParallel::Worker` for performing chunks of Monte Carlo simulations in parallel, simulating (in general) only a subset `subCols` of the variables.
```{Rcpp mcMatRcppParallel, eval=FALSE}
```
The parallel-engine nature of the `yarn2` generator ensures the parallel simulation is playing fair, i.e. is consistent with the sequential simulation.
```{r fullMatExampleRcppParallel}
```
Similarly, we can achieve a consistent parallel simulation of a subset of the variables only.
```{r subMatExampleRcppParallel}
```
```{r, echo=FALSE}
knitr::kable(cbind.data.frame(M = M, Sp = Sp), row.names = TRUE)
```


## Consistent sub-simulaton in R

We rely on the TRNG engines exposed to R by package rTRNG as reference classes.
```{r}
library(rTRNG)
```
We define functions `mcMatR` and `mcSubMatR` performing the full sequential Monte Carlo simulation and the sub-simulation, respectively.
```{r mcMatR}
```
```{r mcSubMatR}
```
As for the Rcpp case above, the behavior of `mcSubMatR` is consistent with the full sequential simulation.
```{r subMatExampleR}
```
```{r, echo=FALSE}
knitr::kable(cbind.data.frame(M = M, S = S), row.names = TRUE)
```
