---
title: "Using rTRNG"
author: "Riccardo Porreca, Roland Schmid"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using rTRNG}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!---
https://rpubs.com/Thell/knitRcpp
-->
## Introduction

**[TRNG](https://numbercrunch.de/trng/)** (Tina's Random Number Generator) is a
state of the art C++ pseudo-random number generator library for sequential and
parallel Monte Carlo simulations. It provides a variety of random number engines
(pseudo-random number generators) and distributions. In particular, *parallel*
random number engines provided by TRNG can be manipulated by `jump` and `split`
operations. These allow to `jump` ahead by an arbitrary number of steps and to
`split` a sequence into any desired sub-sequence(s), thus enabling techniques
such as *block-splitting* and *leapfrogging* suitable to parallel algorithms.

Package **rTRNG** provides the R users with access to the functionality of the
underlying TRNG C++ library, both in R and as part of other projects combining R
with C++.


### TRNG.Random

The TRNG.Random functionality (see `?TRNG.Random`) provides a base-R-like access
to TRNG random number engines by setting and manipulating the current engine in 
use.
```{r TRNG.Random}
library(rTRNG)
TRNGkind("yarn2")
TRNGseed(117)
TRNGjump(5) # advance by 5 the internal state
TRNGsplit(3, 2) # subsequence: one element every 3 starting from the 2nd
```
Random variates from the current engine are then generated using functions
`r<dist>_trng`, e.g. `runif_trng` for the uniform distribution.
```{r runif_trng}
x <- runif_trng(10)
x
```

### TRNG.Engine

Random number engines can be explicitly created and manipulated using reference 
objects from a number of classes (see `?TRNG.Engine`), e.g. `yarn2`.

```{r TRNG.Engine}
rng <- yarn2$new()
rng$seed(117)
# alternative: rng <- yarn2$new(117) 
rng$jump(5) 
rng$split(3, 2) 
```
The engine object is then passed as `engine` argument of any `r<dist>_trng`
function.
```{r runif_trng-engine}
x <- runif_trng(10, engine = rng)
x
```

### Parallel generation

The parallel nature of TRNG random number engines allow multi-threaded
generation of random variates, with guaranteed equivalence to a
purely-sequential generation. Parallel generation is available in `r<dist>_trng`
with argument `parallelGrain > 0` and relies on `RcppParallel`, where the number
of parallel threads is controlled via `RcppParallel::setThreadOptions`.
```{r ex-parallelGrain}
TRNGseed(117)
RcppParallel::setThreadOptions(numThreads = 2)
x_parallel <- runif_trng(1e5, parallelGrain = 100)
TRNGseed(117)
x_serial <- runif_trng(1e5)
identical(x_serial, x_parallel)
```


### Standalone C++

C++ code using the C++ TRNG library and header shipped with **rTRNG** can easily
be sourced via `Rcpp::sourceCpp`, with the `Rcpp::depends` attribute taking care
of linking against the library.

```{Rcpp sourceCpp}
// [[Rcpp::depends(rTRNG)]]
#include <Rcpp.h>
#include <trng/yarn2.hpp>
#include <trng/uniform_dist.hpp>
using namespace Rcpp;
using namespace trng;

// [[Rcpp::export]]
NumericVector exampleCpp() {
  yarn2 rng;
  rng.seed(117);
  // alternative: yarn2 rng(117); 
  rng.jump(5);
  rng.split(3, 1); // note the C++ 0-based index for the subsequence
  NumericVector x(10);
  uniform_dist<> unif(0, 1);
  for (unsigned int i = 0; i < 10; i++) {
    x[i] = unif(rng);
  }
  return x;
}
/*** R
exampleCpp()
*/
```
```{r sourceCpp-R, echo=FALSE}
exampleCpp()
```

