---
title: "Using rTRNG"
author: "Riccardo Porreca, Roland Schmid"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using rTRNG}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!---
https://rpubs.com/Thell/knitRcpp
-->
## Introduction


### TRNG.Random

The TRNG.Random functionality (see `?TRNG.Random`) provides a base-R-like access
to TRNG random number engines by setting and manipulating the current engine in
use.
```{r}
library(rTRNG)
TRNGkind("yarn2")
TRNGset.seed(117)
TRNGjump(5)
TRNGsplit(3, 2)
```
Random variates from the current engine are then generated using functions
`r<dist>_trng`, e.g. `runif_trng` for the uniform distribution.
```{r}
x <- runif_trng(10)
x
```

### TRNG.Engine

Random number engines can be explicitly created and manipulated using reference
objects from a reference objects number of classes (see `?TRNG.Engine`).

```{r}
rng <- yarn2$new()
rng$seed(117)
# alternative: rng <- yarn2$new(117) 
rng$jump(5)
rng$split(3, 2)
```
The engine object is then passed as `engine` argument of any `r<dist>_trng`
function.
```{r}
x <- runif_trng(10, engine = rng)
x
```

### sourceCpp

C++ code using directly the C++ TRNG library and header shipped with **rTRNG**
can easily sourced via `sourceCpp`, with `Rcpp::depends` attribute taking care
of linking against the library.

```{r engine = 'Rcpp'}
// [[Rcpp::depends(rTRNG)]]
#include <Rcpp.h>
#include <trng/yarn2.hpp>
#include <trng/uniform_dist.hpp>
using namespace Rcpp;
using namespace trng;

// [[Rcpp::export]]
NumericVector exampleCpp() {
  yarn2 rng;
  rng.seed(117);
  // alternative: yarn2 rng(117); 
  rng.jump(5);
  rng.split(3, 1); // note the C++ 0-based index for the subsequence
  NumericVector x(10);
  uniform_dist<> unif(0, 1);
  for (int i = 0; i < 10; i++) {
    x[i] = unif(rng);
  }
  return x;
}
/*** R
exampleCpp()
*/
```
```{r, echo=FALSE}
exampleCpp()
```

